Combining checkpointing and data compression\\ for high frequency seismic imaging

### 2018-10-03 paper/paper.tex
Expanding related work

Various strategies are routinely employed to reduce this memory requirement, for example 
Domain decomposition (TODO cite) where the domain of the simulation is distributed over
multiple nodes so as to fit within the memory available on each node. This introduces the overhead
of network communication and might not efficiently use the computational resources on each node.
Another common strategy, which is specific to the domain of FWI, is where only the boundaries of
the domain are stored at each timestep and the rest of the wavefield is reconstructed from the
boundaries when required (TODO CITE). Another commonly used strategy that is applied across
a much wider domain of problems is checkpointing. In this strategy, a subset of the data from the
forward pass is stored (and the rest discarded). The discarded data is recomputed when required
by rerunning the forward pass. Starting from Revolve (TODO cite), which is a provably optimal
scheduling algorithm to implement this strategy, various studies have used and extended this
technique to various scenarios (TODO cite other checkpointing papers).

@@

\section{Related Work}
\cite{cyr2015towards}

\cite{dalmau2014lossy}

@@

TODO cite SZ (Frank Cappello)
\section{Test case using Devito and pyRevolve}
We use Devito \cite{devito-compiler} \cite{devito-api} to setup wave equation solvers for forward and
adjoint modes. Devito is a domain-specific language to enable the rapid development of finite-difference
solvers from a high-level description of the partial differential equation. The simplest version of the wave
equation is the acoustic:
\begin{equation}
m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x)
\end{equation}
where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness with $c(x)$ the spatially dependent speed of sound, 
$u(t, x)$ is the wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is the source.
Some of the kernels used in the performance measurements corresponded to a more physically representative
but complex version of this equation called Tilted Transverse Isotropy (TTI) \cite{zhang2011stable}. We leave the 
equation itself out of this paper for brevity.

The values of $m(x)$ were derived from the SEAM Overthrust (TODO Cite SEAM) model which was over a grid
of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
was TODO in space and TODO in time. This was run for TODO timesteps, and the final field stored as
representative of a typical snapshot. Figure \ref{fig:uncompressed} shows this stored field that was used for the compression tests.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/uncompressed.png}
\end{center}
\caption{Cross-section of the wavefield used as a reference sample for compression and decompression}
\label{fig:uncompressed}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/decompressed-t-8.png}
\end{center}
\caption{Cross-section of the wavefield recovered after compression and decompression using the fixed-tolerance mode}
\label{fig:decompressed}
\end{figure}

@@

>>>>

Various strategies are routinely employed to reduce this memory requirement, for example 
Domain decomposition (TODO cite) where the domain of the simulation is distributed over
multiple nodes so as to fit within the memory available on each node. This introduces the overhead
of network communication and might not efficiently use the computational resources on each node.
Another common strategy, which is specific to the domain of FWI, is where only the boundaries of
the domain are stored at each timestep and the rest of the wavefield is reconstructed from the
boundaries when required (TODO CITE). Another commonly used strategy that is applied across
a much wider domain of problems is checkpointing. In this strategy, a subset of the data from the
forward pass is stored (and the rest discarded). The discarded data is recomputed when required
by rerunning the forward pass. Starting from Revolve (TODO cite), which was a provably optimal
scheduling algorithm to implement this strategy, various studies have used and extended this
technique to various scenarios (TODO cite other checkpointing papers).

@@

\section{Related Work}
\cite{cyr2015towards} studied compression for adjoint-based optimisation in the context of the Navier-Stokes
equation, focussing on the scenario where the entire forward field could be stored in memory after
compression. They looked at Revolve as an alternative to compression.

\cite{dalmau2014lossy} also looked at compression as an alternative to Revolve, although in the same context
as this paper - seismic imaging. Their implementation was, however, storing the compressed data to disk so a
lot of their benefits came from more efficient utilization of the low disk bandwidth. Their focus was on developing
and fine-tuning a domain-specific compression algorithm for seismic imaging which was based on discrete
cosine transforms (DCT).

@@

TODO cite SZ (Frank Cappello)
TODO Jan check whether SZ supports 4D - it does support up to 4D. 
\section{Test case using Devito and pyRevolve}
We use Devito \cite{devito-compiler} \cite{devito-api} to setup wave equation solvers for forward and
adjoint modes. Devito is a domain-specific language to enable the rapid development of finite-difference
solvers from a high-level description of the partial differential equation. The simplest version of the wave
equation is the acoustic:
\begin{equation}
m(x)\frac{\partial^2 u(t, x)}{\partial t^2} - \Laplace u(t, x) = q(t, x)
\end{equation}
where $m(x) = \frac{1}{c^2(x)}$ is the squared slowness with $c(x)$ the spatially dependent speed of sound, 
$u(t, x)$ is the wavefield, $\Laplace u(t, x)$ denotes the laplacian of the wavefield and $q(t, x)$ is the source.
Some of the kernels used in the performance measurements corresponded to a more physically representative
but complex version of this equation called Tilted Transverse Isotropy (TTI) \cite{zhang2011stable}. We leave the 
equation itself out of this paper for brevity.

TODO expand to include FWI objective function and explain that the gradient is computed using the adjoint-state method

The values of $m(x)$ were derived from the SEAM Overthrust (TODO Cite SEAM) model which was over a grid
of $287 \times 881 \times 881$ points, including an absorbing layer of 40 points on each side. The grid spacing
was 25m in space and TODO in time. This was run for 2500 timesteps, and the final field stored as
representative of a typical snapshot. Figure \ref{fig:uncompressed} shows this stored field that was used for the compression tests.

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/uncompressed.pdf}
\end{center}
\caption{Cross-section of the wavefield used as a reference sample for compression and decompression}
\label{fig:uncompressed}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=0.8\linewidth]{images/errors.pdf}
\end{center}
\caption{Cross-section of the field that shows errors introduced during compression and decompression using the fixed-tolerance mode}
\label{fig:decompressed_error}
\end{figure}

@@

More than 3D fields (use time correlation)

Parallel decompression
